Nix for Polyglot, Reproducible Data Science Workflows

### 2025-10-29 paper.qmd
Update paper.qmd

---
title: "Nix for Polyglot, Reproducible Data Science Workflows"
format:
    jss-pdf:
        keep-tex: true
    jss-html:
        embed-resources: true
        toc: true
author:
  - name: Bruno Rodrigues
    affiliations:
      - name: Ministry of Research and Higher education, Luxembourg
        department: Department of Statistics
        address: 18, Montée de la Pétrusse
        city: Luxembourg
        country: Luxembourg
        postal-code: 2327
      - Journal of Statistical Software
    orcid: 0000-0002-3211-3689
    email: bruno@brodrigues.co
    url: https://www.brodrigues.co
  - name: Philipp Baumann
    affiliations:
      - name: Alliance SwissPass, Switzerland
        department: Data and Analytics
        address: 7, Länggassstrasse
        city: Bern
        country: Switzerland
        postal-code: 3012
    orcid: 0000-0002-3194-8975
    email: baumann-philipp@protonmail.com
abstract: |
  Reproducible analysis requires more than clean, well-documented code; it
  demands the precise management of software dependencies, computational
  environments, and workflow execution. While researchers can combine tools like
  [Docker]{.pkg}, [renv]{.pkg} for [R]{.proglang} packages (or [uv]{.pkg} for
  [Python]{.proglang} packages) and [Make]{.pkg} to address these needs, the
  resulting toolchain is often complex and fragile, requiring the coordination
  of multiple disparate systems. This paper introduces a unified framework built
  on the [Nix]{.pkg} package manager, made accessible through two [R]{.proglang}
  packages. The first, [rix]{.pkg}, generates declarative [Nix]{.pkg}
  expressions to define version-pinned, reproducible environments that encompass
  [R]{.proglang}, [Python]{.proglang}, [Julia]{.proglang}, and system-level
  dependencies. The second, [rixpress]{.pkg} (and its [Python]{.proglang} port,
  [ryxpress]{.pkg}), leverages these environments to orchestrate polyglot
  pipelines where each computational step runs in a hermetically sealed,
  language-specific environment with automatic caching. By treating the entire
  computational environment as code, this integrated approach ensures a high
  degree of reproducibility, supports collaboration across heterogeneous
  systems, and helps guarantee that analyses remain executable long into the
  future.

>>>>

---
title: "Nix for Polyglot, Reproducible Data Science Workflows"
format:
    jss-pdf:
        keep-tex: true
    jss-html:
        embed-resources: true
        toc: true
author:
  - name: Bruno Rodrigues
    affiliations:
      - name: Ministry of Research and Higher education, Luxembourg
        department: Department of Statistics
        address: 18, Montée de la Pétrusse
        city: Luxembourg
        country: Luxembourg
        postal-code: 2327
    orcid: 0000-0002-3211-3689
    email: bruno@brodrigues.co
    url: https://www.brodrigues.co
  - name: Philipp Baumann
    affiliations:
      - name: Alliance SwissPass, Switzerland
        department: Data and Analytics
        address: 7, Länggassstrasse
        city: Bern
        country: Switzerland
        postal-code: 3012
    orcid: 0000-0002-3194-8975
    email: baumann-philipp@protonmail.com
abstract: |
  Reproducible analysis requires more than clean, well-documented code; it
  demands the precise management of software dependencies, computational
  environments, and workflow execution. While researchers can combine tools like
  [Docker]{.pkg}, [renv]{.pkg} for [R]{.proglang} packages (or [uv]{.pkg} for
  [Python]{.proglang} packages) and [Make]{.pkg} to address these needs, the
  resulting toolchain is often complex and fragile, requiring the coordination
  of multiple disparate systems. This paper introduces a unified framework built
  on the [Nix]{.pkg} package manager, made accessible through two [R]{.proglang}
  packages. The first, [rix]{.pkg}, generates declarative [Nix]{.pkg}
  expressions to define version-pinned, reproducible environments that encompass
  [R]{.proglang}, [Python]{.proglang}, [Julia]{.proglang}, and system-level
  dependencies. The second, [rixpress]{.pkg} (and its [Python]{.proglang} port,
  [ryxpress]{.pkg}), leverages these environments to orchestrate polyglot
  pipelines where each computational step runs in a hermetically sealed,
  language-specific environment with automatic caching. By treating the entire
  computational environment as code, this integrated approach ensures a high
  degree of reproducibility, supports collaboration across heterogeneous
  systems, and helps guarantee that analyses remain executable long into the
  future.
